Basic Model formulation

### 2015-01-23 documentation/model_formulation.Rmd
model development, accelerating global growth

This document is supposed to become an interactive model representation of the pair-approximation model.

@@

$$ \frac{dV}{dt} = G(V) - C(V)H $$

with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, $H$. Formally, intrinsic mortality could be included in both terms. In the formulations of @noy-meir75 and @vandekoppel97 it is assumed to be covered by the logistic growth function.

@@

defplot <- function(...) plot(..., 
            type = "l", 
            bty = "l", cex = 0.7, las = 1, 
            family = "Times",
            xlim = c(0,1), ylim = c(0,0.2), 
            xaxs = "i", yaxs = "i", 
            xaxp = c(0,1,2), yaxp = c(0,0.2,2), 
            xlab = "vegetation cover")

@@

```{r, echo=FALSE}
G <- function(P, r = 0.5, K = 0.9) r*P*(1-P/K)

@@

```{r, echo=FALSE}
G <- function(P, r = 0.5, K = 0.9) r*P*(1-P/K)

@@

#### decelerating growth at low cover

In arid ecosystems, growth would be decelerated at very low cover, because retention of water and organic matter in the site is fundamentally reduced. This postive feed-back (low cover reduces growth, leading to even lower cover) was discussed by @vandekoppel97 and @rietkerk97 to explain bistability in arid ecosystems.

>>>>

This document is supposed to develop a model representation of the global mean field as well as pair-approximations of local feed back mechanisms.

@@

$$ \frac{dV}{dt} = G(V) - C(V)g $$

with growth function $G$ and a function $C$ which represents consumption by a herbivore multiplied by the density of herbivores, the livestock rate, $g$. Formally, intrinsic mortality could be included in both terms. In the formulations of @noy-meir75 and @vandekoppel97 it is assumed to be covered by the logistic growth function.

@@

defplot <- function(...) plot(..., 
            type = "l", 
            bty = "l", cex = 0.7, las = 1, 
            family = "Times",
            xlim = c(0,1), ylim = c(0,1), 
            xaxs = "i", yaxs = "i", 
            xaxp = c(0,1,2), yaxp = c(0,1,2), 
            xlab = "vegetation cover")

@@

G <- function(P, r = 2, K = 0.9) r*P*(1-P/K)

@@

```{r, echo=FALSE}
G <- function(P, r = 2, K = 0.9) r*P*(1-P/K)

@@

#### accelerating growth

In arid ecosystems, growth would be low at very low cover and accelerate as cover increases, because retention of water and organic matter in the site is fundamentally reduced. This postive feed-back (low cover reduces growth, leading to even lower cover) was discussed by @vandekoppel97 and @rietkerk97 to explain bistability in arid ecosystems.

Starting out from the logistic growth term, we now additionally assume that $r$ is accelerated with cover,

$$ r = r_0 * V^a $$

$$ G(V) = r_0 V^{a+1} (1-V/K) \,,$$

r <- function(P, r_0 = 2, a = 2) r_0*P^a

defplot(rho, r(rho), ylab = "growth")

G <- function(P, r_0 = 2, a = 2, K = 0.9) r(P, r_0, a)*P*(1-P/K)
